---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

As the instructions for this project indicate, this document does not need to include the code related to downloading the data. Instead, I'm going to assume you have [this data]("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip") saved in your working directory.

### Unzip, Load, and Summarize

The data for this project comes in a zip file, but we'll want to extract that in
order to load it into R.

```{r unzip,echo=TRUE}
unzip("week2.zip")
```

This will drop activity.csv into your working directory, so next we'll load the 
data into R:

```{r load}
activity<-read.csv("activity.csv",colClasses = c("integer","Date","integer"))
```

And we'll just want to take a peak at what we've got to work with

```{r summary}
summary(activity)
```

And of course, this lets me know that I've got a lot of NAs to deal with, but we
will get to that later. For now, let's create another dataset where NA values 
are all removed:

```{r omit NAs}
act_clean<-na.omit(activity)
```


## What is mean total number of steps taken per day?

First we need to aggregate total number of steps per day:
```{r aggregate}
daily<-setNames(
      aggregate(
            act_clean$steps,
            by = list(act_clean$date),
            FUN=sum
            ),
      c("Date","Steps")
      )
```

And then we'll create a histogram using ggplot2:
```{r histogram}
qplot(daily$Steps,geom="histogram",binwidth=2000)
```

I wonder what the mean and median steps are?
```{r facts}
mean(daily$Steps)
median(daily$Steps)
```

I love it when median and mean are that close together.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
