---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Loading and preprocessing the data

As the instructions for this project indicate, this document does not need to include the code related to downloading the data. Instead, I'm going to assume you have [this data]("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip") saved in your working directory.

### Unzip, Load, and Summarize

The data for this project comes in a zip file, but we'll want to extract that in
order to load it into R.

```{r unzip,echo=TRUE}
unzip("week2.zip")
```

This will drop activity.csv into your working directory, so next we'll load the 
data into R:

```{r load}
activity<-read.csv("activity.csv",colClasses = c("integer","Date","integer"))
```

And we'll just want to take a peak at what we've got to work with

```{r summary}
summary(activity)
```

And of course, this lets me know that I've got a lot of NAs to deal with, but we
will get to that later. For now, let's create another dataset where NA values 
are all removed:

```{r omit NAs}
act_clean<-na.omit(activity)
```


## What is mean total number of steps taken per day?

First we need to aggregate total number of steps per day:
```{r aggregate}
daily<-setNames(
      aggregate(
            act_clean$steps,
            by = list(act_clean$date),
            FUN=sum
            ),
      c("Date","Steps")
      )
```

And then we'll create a histogram using ggplot2:
```{r histogram}
qplot(daily$Steps,geom="histogram",binwidth=2000)
```

I wonder what the mean and median steps are?
```{r facts}
mean(daily$Steps)
median(daily$Steps)
```

I love it when median and mean are that close together.


And what interval has the highest average step count?

```{r max steps}
daily_steps[daily_steps$mean_steps==max(daily_steps$mean_steps),]
```

Looks like interval 835 wins with 206.1698 steps on average

## What is the average daily activity pattern?

So what does the average daily pattern look like? First, let's make a new
dataset that aggregates the average number of steps per 5 minute interval:

```{r pattern aggregate}
daily_steps<-setNames(aggregate(
            act_clean$steps, 
            by = list(act_clean$interval), 
            FUN="mean"
            ), 
      c("interval","mean_steps"))

head(daily_steps)
```

Great. Now let's plot that!

```{r daily avg steps}
ggplot(daily_steps,aes(interval,mean_steps))+
      geom_line(color = "sky blue",size=1)+
      theme_bw()
```

We can start to see a pretty clear pattern. Looks like the test subject(s) tend
to take a morning walk of some kind. 

## Imputing missing values

Remeber how earlier we said there were a bunch of NAs in our 'steps' data? 
Well, there are. In fact, there are this many:
```{r}
sum(is.na(activity$steps))
```

I wonder if ignoring all of those NAs has distorted my data to some degree. So 
let's do something about them and see what happens:

```{r imputing}
act2<-activity
act2$steps[is.na(act2$steps)]<-mean(act2$steps,na.rm = TRUE)
summary(act2$steps)
```

Great, with no NAs left in my new 'act2' dataset, let's look at how the data may
have changed.

```{r data comparison}
daily2<-setNames(
      aggregate(
            act2$steps, 
            by = list(act2$date), 
            FUN=sum
            ), 
      c("Date","Steps"))

qplot(daily2$Steps,geom="histogram",binwidth=2000)

mean(daily2$Steps)
median(daily2$Steps)

```

So, we can see that the shape of the histogram has shifted a bit and that the 
median value is much more clear. Now, the mean and median are identical, and 
because of the way that I imputed the NAs, the new median and mean match the 
original mean of the dataset.

## Are there differences in activity patterns between weekdays and weekends?

We'd like to see if there is a difference in weekdays vs weekends insofar as 
daily activity goes. So first, we'll need to create a new variable that identifies
if a day falls on a weekday or a weekend.

```{r weekpart}
act2$weekpart<-as.factor(
      ifelse(
            weekdays(act2$date) %in%
                  c("Saturday","Sunday"),
            "Weekend",
            "Weekday"
      )
)
```

Then we can plot activity on weekend days vs weekdays and see what difference
may exist.

```{r weekpart visual}
weekpart.daily.steps<-setNames(
      aggregate(
            act2$steps,by=list(act2$interval,act2$weekpart),
            FUN="mean"
      ),
      c("interval","weekpart","meansteps")
)

ggplot(weekpart.daily.steps,aes(interval,meansteps))+
      geom_line()+
      ylab("Number of Steps")+
      xlab("Interval")+
      theme_bw()+
      facet_wrap(~weekpart,nrow=2,ncol=1)
      
      
```

There's definitely less of a spike on the weekends in the early morning (maybe 
we are missing a walk from the parking lot to the office building?) and more
midday steps during the weekend, which may indicate our subject(s) have a more
stationary desk job during the week.
